### 2022-05-12 meanNetPayoffs.tex
updated figures and analysis text

\begin{figure}
  \caption{Dependence of average net payoff normalized by lifespan,
    $\meanpi/L$, at the
    end of last time step of the simulation on environmental
    variability $u$ for several values of low payoff $\pilow$, number of
    behaviors $B$, and lifespan $L$. Dashed lines mark expected 
    payoff from individual learning $\langle \pi_I \rangle$ calculated
    for each parameter setting (not sensitive to $u$). 
    Note that when the maximum $L$ is 
    8, the y-axis runs from 0 to 8, and when the maximum $L$ is 20, the y-axis
    runs from 0 to 20. Average over 1000 trials.} 
  \label{fig:meanNetPayoffs}
  \centering
  \begin{subfigure}[t]{0.08\textwidth}
    \begin{tikzpicture}
      \draw[->,thick] (-.1,0)--(1,0) node[above]
        {$B$};
      \draw[->,thick] (0,.1)--(0,-1) node[below]
        {$\pilow$};
    \end{tikzpicture}

  \end{subfigure}%
  \begin{subfigure}[t]{0.3\textwidth}
    \centering
    {\large $B = 2$} 
    \begin{rotate}{90}
      {\parbox{4cm}{
          \centering
          \vspace{-3.5em}{\large $ \pilow = 0.1$} \\
          {\begin{rotate}{-90}{\large $\frac{\meanpi}{L}$}\end{rotate}}
      }}
    \end{rotate}%
    \includegraphics[width=1.1\textwidth]{
      {Figures/mean_prev_net_payoff_over_u_lowpayoff=0.1_nbehaviors=2.pdf}
    }
    \centering
    \begin{rotate}{90}
      {\parbox{4cm}{
          \centering
          \vspace{-3.5em} {\large$ \pilow = 0.45$} \\
          {\begin{rotate}{-90}{\large $\frac{\meanpi}{L}$}\hspace{1em}\end{rotate}}
      }}
    \end{rotate}%
    \includegraphics[width=1.1\textwidth]{
      {Figures/mean_prev_net_payoff_over_u_lowpayoff=0.45_nbehaviors=2.pdf}
    }
    \begin{rotate}{90}
      {\parbox{4cm}{
          \centering
          \vspace{-3.5em} {\large $ \pilow = 0.8$} \\
          {\begin{rotate}{-90}{\large $\frac{\meanpi}{L}$}\end{rotate}}
      }}
    \end{rotate}%
    \includegraphics[width=1.1\textwidth]{
      {Figures/mean_prev_net_payoff_over_u_lowpayoff=0.8_nbehaviors=2.pdf}
    }\\[-.5em]
    {\large $\quad u$}
  \end{subfigure}%
  \begin{subfigure}[t]{0.3\textwidth}
    {\large $B = 4$}
    \includegraphics[width=1.1\textwidth]{
      {Figures/mean_prev_net_payoff_over_u_lowpayoff=0.1_nbehaviors=4.pdf}
    } \\
    \centering
    \includegraphics[width=1.1\textwidth]{
      {Figures/mean_prev_net_payoff_over_u_lowpayoff=0.45_nbehaviors=4.pdf}
    } \\
    \includegraphics[width=1.1\textwidth]{
      {Figures/mean_prev_net_payoff_over_u_lowpayoff=0.8_nbehaviors=4.pdf}
    }\\[-.5em]
    {\large{$\quad u$}}
  \end{subfigure}
  \begin{subfigure}[t]{0.3\textwidth}
    \centering
    {\large $B = 10$}
    \includegraphics[width=1.1\textwidth]{
      {Figures/mean_prev_net_payoff_over_u_lowpayoff=0.1_nbehaviors=10.pdf}
    } \\
    \centering
    \includegraphics[width=1.1\textwidth]{
      {Figures/mean_prev_net_payoff_over_u_lowpayoff=0.45_nbehaviors=10.pdf}
    } \\
    \includegraphics[width=1.1\textwidth]{
      {Figures/mean_prev_net_payoff_over_u_lowpayoff=0.8_nbehaviors=10.pdf}
    } \\[-.5em]
    {\large $\quad u$}
  \end{subfigure}

\end{figure}